### 2019-06-12 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Commentaire Christophe Ch4.1 et 4.2
Commentaires de Xtof pour le Ch4

In this section, we present a demonstration showed at the International Conference on Telecommunication (ICT) in June $2018$ \cite{Besson2018ICT,Besson2019WCNC}, implementing a proof-of-concept (PoC) of the model introduced in Section~\ref{sec:4:firstModel}.

@@

In our demo, we consider a simple wireless network, consisting of one gateway (\ie, radio access point), and a certain interfering background traffic, assumed to be stationary (\emph{i.i.d.}), which is generated by end-devices communicating in other networks.
Some dynamic intelligent devices (end-user or autonomous devices) try to communicate with the gateway, with a low-overhead protocol. This communication can be done in different channels which are also shared by devices using other networks.
Once the gateway receives a packet transmitted by a dynamic device in one channel (\ie, if no collision occurred), it transmits back to it an acknowledgement in the same channel, after a fixed-time delay, as it is done in the LoRaWAN standard.
This \emph{Ack} allows the device to learn about the channel quality (\ie, mean availability) and thus, to use learning algorithms for the purpose of best channel selection.

We can generate scenarios with different parameters (number of channels, interfering traffic, etc) in order to evaluate the performance of learning in various settings.
Moreover, we compare the performance of learning strategies with that of the random uniform access to channels, which is the current state-of-the-art of commercial LPWAN solutions \cite{Raza17}.
%
This allows to check that in case of uniform traffic, when there is nothing to learn, the intelligent devices at least do not reduce their successful communication rate in comparison to the naive devices.
This also shows that in case of non-uniform stationary traffic, MAB learning algorithms indeed help to increase the global efficiency of the network by improving the success rate of the intelligent devices.
%
The benefits are twice and of primary importance for IoT networks:
the proposed approach can mitigate RF collisions,
and enhance intelligent device battery lifetime if they do retransmissions.

@@

\textbf{Physical layer and protocol.}
%
We implement a simple PHY/MAC layers solution, in order to demonstrate the possibility of improvement of the performance of IoT communications in unlicensed bands. We could have used any physical layer and any ALOHA-based protocol.
We choose to implement our own physical layer and protocol, for both clarity and conciseness, and because developing a complete IoT network protocol stack is no more a research work and would have fall outside of the scope of this thesis.

Regarding the physical layer, we consider a QPSK constellation (\emph{Quadrature Phase-Shift Keying}). Moreover, we use simplified packets composed of two parts.
The first part is the \emph{preamble} which is used for the purpose of synchronization (phase correction).
Then, we have the \emph{index} of the user, which is a sequence of QPSK symbols.
For example, this index can be a simple QPSK symbol ($\pm1\pm1j$), or a longer sequence if the PoC have to consider more users.
With $\ell$ QPSK symbols, we can indeed fit at most $4^{\ell}/2 = 2^{2\ell-1}$ devices, and not $4^{\ell}$ due to the use of a conjugate index to send back acknowledgements from the gateway
Once the gateway receives an uplink packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the uplink packet ($z \mapsto \overline{z}$, \emph{e.g.}, $1+j \mapsto 1-j$).
Thanks to this index, we can have several devices communicating with the same gateway.
%
In turn, the end-device that receives the acknowledgement demodulates it, and checks if the index is the conjugate of its own index.
In this case, the \emph{Ack} was sent for him, and it knows that its packet has been received and decoded correctly by the gateway.

@@

\paragraph{Equipment.}
% Boards and versions
We use USRP N210 boards \cite{USRPDocumentation}, from Ettus Research (National Instrument),
with version $4$ of their FPGA system and version $5.1$ of the RBX system.
As illustrated in Figure~\ref{fig:42:our_demo}, our implementation is composed of at least $3$ USRP.
The gateway, a USRP which emulates the interfering traffic, and at least one dynamic device, each of them having its own USRP and its own learning algorithm.
% Cables and laptop
% Osef

@@

The boards have their own power supply, and are all connected to a local Ethernet switch, itself connected to a single laptop, running GNU/Linux and Ubuntu.
The pictures in Figure~\ref{fig:42:photosSCEETestBed} shows the testbed used for these experiments.
% Octoclock
To ease the synchronization in both time and frequency between the boards representing the dynamic devices and the gateway, we use an Octoclock \cite{OctoclockProduct}, also a product of Ettus Research,
% https://www.ettus.com/product/details/OctoClock-G
and coaxial cables connecting every platform to the Octoclock for time (PPS) and frequency synchronization, but this is not mandatory.

@@

\begin{enumerate}[leftmargin=6mm]
    \item[(1)]
The first part is the interface of the IoT traffic generator, where we see the traffic generated by this USRP, presented in a waterfall view in the time vs frequency domain.
Messages of the random traffic are shorter in time by purpose (they could be coming from othe rIoT standard), in order to distinguish them from an intelligent device traffic on the ``waterfall'' visualizations of the traffic.

    \item[(2)]
The second part is the interface of the intelligent device which is made of four parts.
At the top left, we observe the constellation of the transmitted packet \emph{(a)}.
At the bottom left, we have a time/frequency view of the lasts packets transmitted by the device \emph{(b)}.
We can see, in this view that the device transmitted its last $9$ packets in the two best channels (channel $\#3$ and $\#4$).
Then, at the top right of this interface \emph{(c)}, we can see the traffic observed by this device, where we have the interfering traffic (\textcolor{darkgreen}{green}), the uplink packets transmitted by this device (\textcolor{darkred}{red}) and the acknowledgements sent by the gateway (\textcolor{darkblue}{blue}).
Colors in the ``waterfall'' represent the RF power level received at the device antenna. Hot colors are for closer elements, as for instance the device Tx antenna (reception) is close to the device Rx antenna (emission), and consequently for the device waterfall these signals are colored in \textcolor{darkred}{red} (see graph \emph{(c)} in part (2)).
Finally, at the bottom right \emph{(d)}, we have four histograms showing the performance indicators of the chosen MAB algorithm (number of transmissions, number of successful transmissions, UCB indexes and success rates, in each channel).

    \item[(3)]
The last part is the interface of the gateway, where we can see the traffic observed by the gateway \emph{(a)} and the channels in which the last acknowledgements have been sent \emph{(b)}.
\end{enumerate}

% ----------------------------------------------------------------------
\subsection{Experimental results}
\label{sub:42:results}
% ----------------------------------------------------------------------
We compare in this PoC the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} (\UCB{} and Thompson sampling) against a uniform access algorithm, that uniformly selects its channel at random.
For one dynamic device, three algorithms are compared by their mean successful communication rates, on a horizon of $T=2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (``\textcolor{green}{TS}'', in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).

@@

In Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} below, we show the results averaged on $10$ repetitions using the same conditions.
%
Each experiment takes about half a day,
as we make devices generate one message every $5$ seconds, in order to artificially speed up the process and with no loss of generality (as we are using real hardware).
Learning can be useful only when there is a large enough difference between ``good'' and ``bad'' channels,
Each device was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $(\mu_i)_i = [15\%, 10\%, 2\%, 1\%]$.
The occupancy rate of a channel, which denotes the mean occupancy, is implemented using the traffic generator, to emulate the presence of $S_i$ static devices with emission probability $p$ (that is, $\mu_i = S_i \times p$ here).
% They did not overlap each other, as
When facing the same stationary background traffic, we see that the learning devices are both very quickly more efficient than the naive uniform device.
We obtain an improvement in terms of successful communication rate from $40\%$ to about $60\%$ in only $100$ communications (about $16\;\mathrm{min}$), and up-to $80\%$ in only $400$ communications.
%
In stationary environments, both the TS and \UCB{} algorithms are very efficient and converge quickly, resulting in a very strong decrease in collisions and failed communication slots. \UCB{} is faster to learn but eventually TS gives a (slightly) better average performance.

>>>>

In this section, we present a demonstration showcased at the International Conference on Telecommunication (ICT) in June $2018$ \cite{Besson2018ICT,Besson2019WCNC}, implementing a proof-of-concept (PoC) of the model introduced in Section~\ref{sec:4:firstModel}.
%
As far as we know, this is the first demonstration of running learning algorithms on the IoT device side, in real radio conditions, as highlighted it in our paper \cite{MoyBesson2019}.

@@

In our demo, we consider a simple wireless network, that reproduces the model of Section~\ref{sec:4:firstModel}, consisting of one gateway (\ie, radio access point), and a certain interfering background traffic, assumed to be stationary (\emph{i.i.d.}), which is generated by end-devices communicating in other networks.
Some dynamic intelligent devices (end-user or autonomous devices) try to communicate with the gateway, with a low-overhead protocol. This communication can be done in different channels which are also shared by devices using other networks.
Once the gateway receives a packet transmitted by a dynamic device in one channel (\ie, if no collision occurred), it transmits back to it an acknowledgement in the same channel, after a fixed-time delay, as it is done in the LoRaWAN standard.
This \emph{Ack} allows the device to learn about the channel quality (\ie, mean availability) and thus, to use learning algorithms for the purpose of best channel selection.

We can generate scenarios with different parameters (number of channels, interfering traffic load on each channel, etc) in order to evaluate the performance of learning in various settings.
Moreover, we compare the performance of learning strategies with that of the random uniform access to channels, which is the current state-of-the-art of commercial LPWAN solutions \cite{Raza17}.
%
This allows to check that in case of uniform traffic, when there is nothing to learn, the intelligent devices at least do not reduce their successful communication rate in comparison to the naive devices.
This also shows that in case of non-uniform stationary traffic, MAB learning algorithms indeed help to increase the global efficiency of the network by improving the success rate of the intelligent devices.
%
The benefits are twice and of primary importance for IoT networks:
the proposed approach can mitigate RF collisions,
and enhance intelligent device battery lifetime if they do retransmissions.

@@

\textbf{Physical layer and protocol.}
%
We implement a simple PHY/MAC layers solution, in order to demonstrate the possibility of improvement of the performance of IoT communications in unlicensed bands. We could have used any physical layer and any ALOHA-based protocol.
We choose to implement our own physical layer and protocol, for both clarity and conciseness, and because developing a complete IoT network protocol stack is no more my research work and would have fall outside of the scope of this thesis.

Regarding the physical layer, we consider a QPSK constellation (\emph{Quadrature Phase-Shift Keying}). Moreover, we use simplified packets composed of two parts.
The first part is the \emph{preamble} which is used for the purpose of synchronization (phase correction).
Then, we have the \emph{index} of the user, which is a sequence of QPSK symbols.
For example, this index can be a simple QPSK symbol ($\pm1\pm1j$), or a sequence of QPSK symols if the PoC has to consider more users.
With $\ell$ QPSK symbols, we can indeed fit at most $4^{\ell}/2 = 2^{2\ell-1}$ devices, and not $4^{\ell}$ due to the use of a conjugate index to send back acknowledgements from the gateway.
Once the gateway receives an uplink packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the uplink packet ($z \mapsto \overline{z}$, \emph{e.g.}, $1+j \mapsto 1-j$).
Thanks to this index, we can have several devices communicating with the same gateway.
%
In turn, the end-device that receives the acknowledgement demodulates it, and checks if the index is the conjugate of its own index.
In this case, the \emph{Ack} was sent for him, and it knows that its packet has been received and decoded correctly by the gateway.

@@

\paragraph{Equipment.}
% Boards and versions
We use USRP N210 boards \cite{USRPDocumentation}, from Ettus Research (National Instrument),
with version $4$ of their FPGA system and version $5.1$ of the RBX system.
As illustrated in Figure~\ref{fig:42:our_demo}, our implementation is composed of at least $3$ USRP:
the gateway,
a traffic generator which emulates the interfering traffic (made by surrounding static devices),
and at least one dynamic device.
Each dynamic device has its own USRP and its own learning algorithm.
% Cables and laptop
% Osef

@@

The boards have their own power supply, and are all connected to a local Ethernet switch, itself connected to a single laptop, running GNU/Linux and Ubuntu.
The pictures in Figure~\ref{fig:42:photosSCEETestBed} show the testbed used for these experiments.
% Octoclock
To ease the synchronization in both time and frequency between the boards representing the dynamic devices and the gateway, we use an Octoclock \cite{OctoclockProduct}, also a product of Ettus Research,
% https://www.ettus.com/product/details/OctoClock-G
and coaxial cables connecting every platform to the Octoclock for time (PPS) and frequency synchronization, but this is not mandatory.

@@

\begin{enumerate}[leftmargin=6mm]
    \item[(1)]
The first part is the interface of the IoT traffic generator, where we see the traffic generated by this USRP, presented in a waterfall view in the time vs frequency domain.
Messages of the random traffic, generated by surrounding static devices, are shorter in time by purpose (they could be coming from other IoT standard), in order to distinguish them from an intelligent device traffic on the ``waterfall'' visualizations of the traffic.

    \item[(2)]
The second part is the interface of the intelligent device which is made of four parts.
At the top left, we observe the constellation of the transmitted packet \emph{(a)}.
At the bottom left, we have a time/frequency view of the lasts packets transmitted by the device \emph{(b)}.
We can see, in this view that the device transmitted its last $9$ packets in channels $\#3$ and $\#4$.
Then, at the top right of this interface \emph{(c)}, we can see the traffic observed by this device, where we have the interfering traffic (\textcolor{darkgreen}{green}), the uplink packets transmitted by this device (\textcolor{darkred}{red}) and the acknowledgements sent by the gateway (\textcolor{darkblue}{blue}).
Colors in the ``waterfall'' represent the RF power level received at the device antenna. Hot colors are for closer elements, as for instance the device Tx antenna (reception) is close to the device Rx antenna (transmission), and consequently for the device waterfall these signals are colored in \textcolor{darkred}{red} (see graph \emph{(c)} in part (2)).
Finally, at the bottom right \emph{(d)}, we have four histograms showing the performance indicators of the chosen MAB algorithm (number of transmissions, number of successful transmissions, UCB indexes and success rates, in each channel).

    \item[(3)]
The last part is the interface of the gateway, where we can see the traffic observed by the gateway \emph{(a)} and the channels in which the last acknowledgements have been sent \emph{(b)}.
The observed traffic is coherent with \emph{(c)} of part (2), as the elements are very close the one from the others on the testbed. Colors may change, as they depend on the exact distance between the different transmitters and receivers.
\end{enumerate}

% ----------------------------------------------------------------------
\subsection{Experimental results}
\label{sub:42:results}
% ----------------------------------------------------------------------
We compare in this PoC the two algorithms described in Section~\ref{sec:2:famousMABalgorithms} (\UCB{} and Thompson sampling) against a uniform access algorithm, that uniformly selects its channel at random.
Note that we could run more algorithms, but with no real added-valued in terms of validation of the proposed learning-based approach, which is our goal.
%
For one dynamic device, three algorithms are compared by their mean successful communication rates, on a horizon of $T=2000$ communication slots, and were using three algorithms: uniform random access (in \textcolor{cyan}{cyan}), Thompson Sampling (``\textcolor{green}{TS}'', in \textcolor{green}{green}) and \UCB{} (in \textcolor{red}{red}).

@@

In Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} below, we show the results averaged on $10$ repetitions using the same conditions.
%
Each experiment has been done on a duration of about half a day,
due to the IoT sporadic transmission mode that we want to respect (like in our model of Section~\ref{sec:4:firstModel}).
However, we make devices generate one message every $5$ seconds, in order to artificially speed up the process and with no loss of generality (as we are using real hardware).
Learning can be useful only when there is a large enough difference between ``good'' and ``bad'' channels,
Each device was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $(\mu_k)_k = [15\%, 10\%, 2\%, 1\%]$.
Note that a maximum occupancy rate of $15\%$ could seem not so high, but indeed it is, because for a pure ALOHA access mode, a naive dynamic device only enjoys about $40\%$ success rate, under such occupancy of the channels (see the ``uniform'' plot in Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}).
%
The occupancy rate of a channel, which denotes the mean occupancy, is implemented using the traffic generator, to emulate the presence of $S_i$ static devices with emission probability $p$ (that is, $\mu_i = S_i \times p$ here).
% They did not overlap each other, as
When facing the same stationary background traffic, we see that the learning devices are both very quickly more efficient than the naive uniform device.
We obtain an improvement in terms of successful communication rate from $40\%$ to about $60\%$ in only $100$ communications (about $16\;\mathrm{min}$), and up-to $80\%$ in only $400$ communications.
%
In stationary environments, both the TS and \UCB{} algorithms are very efficient and converge quickly, resulting in a very strong decrease in collisions and failed communication slots. \UCB{} is faster to learn but eventually TS gives a (slightly) better average performance.
